An examination of the high-order dynamic interactions underlying multi-dimensional timeseries data

### 2018-10-16 paper/main.tex
backing up changes; still working through methods [WIP]

\subsubsection*{Inter-subject dynamic correlations}
Equation~\ref{eqn:timecorr} provides a means of taking a single
observation matrix, $\mathbf{X}$ and estimating the dynamic
correlations from moment to moment, $\mathbf{Y}$.  Suppose that one
has access to a set of multiple observation matrices that reflect the
same phenomenon.  For example, one might collect neuroimaging data
from several experimental participants, as each participant performs
the same task (or sequence of tasks).  Let $\{ mathbf{X}_1,
mathbf{X}_2, ..., mathbf{X}_P\}$ reflect the $T$ by $F$ observation
matrices for each of $P$ participants in an experiment.  We can use
\textit{inter-subject functional connectivity}~\citep[ISFC;
][]{SimoEtal16} to compute the degree of stimulus-driven correlations
reflected in the multi-participant dataset at a given timepoint $t$
using:
\begin{align}
\bar{\mathbf{C}}(t) = M\left(R\left(\frac{1}{2P} \sum_{i=1}^P
  Z\left(Y_i(t)\right)^\mathrm{T} + Z\left(Y_i,(t)\right)\right)\right),
\end{align}
where $M$ extracts and vectorizes the diagonal and upper triangle of a symmetric
matrix, $Z$ is the Fisher $z$-transformation~\citep{Zar10}:
\begin{align}
Z(r) = \frac{\log(1+r) - \log(1-r)}{2}
\end{align}
$R$ is the inverse of $Z$:
\begin{align}
R(z) = \frac{\exp(2z - 1)}{\exp(2z + 1)},
\end{align}
and $\mathbf{Y}_i(t)$ denotes the correlation matrix
(Eqn.~\ref{eqn:corr}) between each column of $\mathbf{X}_i$ and each
column of the average observations from all \textit{other}
participants, $\bar{\mathbf{X}}_{ \backslash i}$:
\begin{align}
  \bar{\mathbf{X}}_{ \backslash i} = R\left(\frac{1}{P-1}\sum_{i \in \backslash i} Z(\mathbf{X}_i\right),
\end{align}
where $ \backslash i$ denotes the set of all participants other than
participant $i$. In this way, the $T$ by $\left( \frac{F^2 - F}{2}
  + F \right)$
matrix $\bar{\mathbf{C}}$ is the time-varying extension of the ISFC
approach developed by \cite{SimoEtal16}.

@@

\subsection*{Timepoint decoding}

Cite HTFA paper, summarize relevant methods.

\subsection*{Prediction}
Use Eqn. 12 from SuperEEG paper, extend notation to treat
$\hat{\mathbf{Y}}_{1...n}(1...(t-1))$ as ``observed'' and use this to
predict $\mathbf{X}(t)$.  Key question: what is the predictive power
added by each new level.

\section*{Results}
\subsubsection*{Synthetic data}
\subsubsection*{Neuroimaging dataset~\citep{SimoEtal16}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}
% multiple timescale representations (a la Hasson group) implies
% first-order network interactions.  higher-order interactions imply
% generalizations between interacting representations (e.g. mirrored
% schema, a la Norman/Baldassano/Hasson).  possibly cite NTB 2013
% science review, using as evidence that this is where the field is
% going (voxels --> patterns (L0) --> interactions (L1) --> higher
% order patterns (L2+).

% related approaches: sliding window, phase-based correlations,
% within-ROI spatial correlations at each timepoint, granger
% causality, other explit models (e.g. virtual brain).

%other applications: molecular interactions (protein folding?),
%diagnosis (e.g. psychiatric disorders as network flow problems--
%gratton work?), social network dynamics (e.g. financial markets,
%social media interactions)

\subsection*{Concluding remarks}
% the universe is complicated and we need scalable approaches to
% studying how the pieces are interacting to make sense of it.  one
% small step for mankind, and so on.

>>>>

\subsubsection*{Inter-subject dynamic correlations}
Equation~\ref{eqn:timecorr} provides a means of taking a single
observation matrix, $\mathbf{X}$ and estimating the dynamic
correlations from moment to moment, $\mathbf{Y}$.  Suppose that one
has access to a set of multiple observation matrices that reflect the
same phenomenon.  For example, one might collect neuroimaging data
from several experimental participants, as each participant performs
the same task (or sequence of tasks).  Let $\{ mathbf{X}_1,
mathbf{X}_2, ..., mathbf{X}_P\}$ reflect the $T$ by $F$ observation
matrices for each of $P$ participants in an experiment.  We can use
\textit{inter-subject functional connectivity}~\citep[ISFC;
][]{SimoEtal16} to compute the degree of stimulus-driven correlations
reflected in the multi-participant dataset at a given timepoint $t$
using:
\begin{align}
\bar{\mathbf{C}}(t) = M\left(R\left(\frac{1}{2P} \sum_{i=1}^P
  Z\left(Y_i(t)\right)^\mathrm{T} + Z\left(Y_i,(t)\right)\right)\right),
\end{align}
where $M$ extracts and vectorizes the diagonal and upper triangle of a symmetric
matrix, $Z$ is the Fisher $z$-transformation~\citep{Zar10}:
\begin{align}
Z(r) = \frac{\log(1+r) - \log(1-r)}{2}
\end{align}
$R$ is the inverse of $Z$:
\begin{align}
R(z) = \frac{\exp(2z - 1)}{\exp(2z + 1)},
\end{align}
and $\mathbf{Y}_i(t)$ denotes the correlation matrix
(Eqn.~\ref{eqn:corr}) between each column of $\mathbf{X}_i$ and each
column of the average observations from all \textit{other}
participants, $\bar{\mathbf{X}}_{ \backslash i}$:
\begin{align}
  \bar{\mathbf{X}}_{ \backslash i} = R\left(\frac{1}{P-1}\sum_{i \in
  \backslash i} Z\left( \mathbf{X}_i \right) \right),
\end{align}
where $ \backslash i$ denotes the set of all participants other than
participant $i$. In this way, the $T$ by $\left( \frac{F^2 - F}{2}
  + F \right)$
matrix $\bar{\mathbf{C}}$ is the time-varying extension of the ISFC
approach developed by \cite{SimoEtal16}.

@@

\subsection*{Evaluation metrics}
We evaluate our approach to extracting dynamic correlations and
higher-order correlations using several metrics detailed next.  First,
we generated synthetic data using known time-varying correlations, and
then we evaluated the fidelity with which Equation~\ref{eqn:timecorr}
could recover those correlations (for synthetic datasets with
different properties, and using different kernels to define the
weights; Fig.~\ref{fig:weights}).  We then turned to a series of
analyses on a (real) neuroimaging dataset where the ground truth
correlations were \textit{not} known.  We evaluated whether the
recovered correlations could be used to accurately label held-out
neuroimaging data with the time at which it was collected, and also
whether the history of correlations (and higher-order correlations)
from time 0 to $t-1$ could be used to predict future activity patterns
at time $t$.  We used these latter evaluations (using timepoint
decoding and predictions of held-out future data) as a proxy for
gauging how much explanatory power the recovered correlations held
with respect to the observed data.

\subsubsection*{Generating synthetic data}
Ramping dataset and block dataset.  Create an example figure for both.

\subsubsection*{Recovery of ground truth parameters from synthetic
  data}
Apply timecorr with a given kernel, then correlate each recovered
correlation matrix with the ground truth.  Explore how recovery varies
with the kernel, kernel parameters, and specific structure of the data
(e.g. slow changes as in the ramping dataset, versus rapid changes as
in the block dataset).

\subsubsection*{Timepoint decoding}
Cite HTFA paper, summarize relevant methods.

\subsubsection*{Predicting held-out future data}
Use Eqn. 12 from SuperEEG paper, extend notation to treat
$\hat{\mathbf{Y}}_{1...n}(1...(t-1))$ as ``observed'' and use this to
predict $\mathbf{X}(t)$.  Key question: what is the predictive power
added by each new level?

\section*{Results}
\subsubsection*{Synthetic data}
Figure: overall timecourse of recovery, also recovery near
event boundaries.

\subsubsection*{Neuroimaging dataset~\citep{SimoEtal16}}
Figures: decoding by level, predictions by level

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}
\begin{itemize}
\item multiple timescale representations (a la Hasson group) implies
first-order network interactions.  higher-order interactions imply
generalizations between interacting representations (e.g. mirrored
schema, a la Norman/Baldassano/Hasson).  possibly cite NTB 2013
science review, using as evidence that this is where the field is
going (voxels $\rightarrow$ patterns (L0) $\rightarrow$ interactions (L1) $\rightarrow$ higher
order patterns (L2$+$).
\item related approaches: sliding window, phase-based correlations,
within-ROI spatial correlations at each timepoint, granger
causality, other explit models (e.g. virtual brain).
\item other applications: molecular interactions (protein folding?),
diagnosis (e.g. psychiatric disorders as network flow problems--
gratton work?), social network dynamics (e.g. financial markets,
social media interactions)
\end{itemize}

\subsection*{Concluding remarks}
the universe is complicated and we need scalable approaches to
studying how the pieces are interacting to make sense of it.  one
small step for mankind, and so on.
